---
title: "`lydemap`: an R package to map <i>Lycorma delicatula</i>"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 3
  html_document:
    toc: TRUE
    toc_depth: 3
    toc_float: true
    theme: united
params:
  show_code: FALSE
author: 
    - "Sebastiano De Bona^[Temple University, sebastiano.debona@temple.edu]"
    - "Matthew R. Helmus^[Temple University, mrhelmus@temple.edu]"
date: "`r format(Sys.time(), '%d %B %Y')`"
---

```{r setting figure files as png, include = F}
# paste this to the above yaml to add authors
#author: 
#    - "Sebastiano De Bona^[Temple University, sebastiano.debona@temple.edu]"
#    - "Matthew R. Helmus^[Temple University, mrhelmus@temple.edu]"
# authors can be removed if necessary

# above, in the YAML parameters, we have defined want_code and 
# export_figures, which will define echo and eval for the code chunks.
# these parameters define whether code is to be displayed
# in the vignettes or no, and whether the figures should be exported as images, respectively.
knitr::opts_chunk$set(dev = "png", dpi = 100, echo = params$show_code)
```

## Introduction

The Spotted lanternfly (*Lycorma delicatula*, White 1841) is an agricultural pest native of China and Southeast Asia, first discovered in the United states in 2014 in Berks County, PA (REFS). Since then, this planthopper has spread throughout the Mid-Atlantic and Midwest regions of the country, threatening the wine and fruit industry and damaging ornamental trees.

Since its first discovery, many sources have collected data on the presence/absence and population density of this species in order to monitor its spread and impact. The `lydemap` package contains two anonymized datasets (at 1 km^2^ and 10 km^2^ resolution) resulting from an effort to combine, organize, and aggregate all available sources of data. In addition, this package contains useful functions to visualize the data within R.

The `lydemap` package was built with the intent to increase accessibility to key data on this species of interest, and to improve reproducibility and consistency of modeling efforts.

We are constantly looking to expand the data sources to have a full representation of SLF's presence and abundance in the US. If you wish to contribute to this effort please contact the package authors.

```{r packages and data, include = params$show_code, message = FALSE, warning = FALSE}
# attaching necessary packages
library(lydemap)
library(sf)
library(tidyverse)
library(tigris)

# factorizing and ordering pop density
lyde <- lyde %>% 
  mutate(slf_density = factor(slf_density, levels = c("Unpopulated",
                                                      "Low",
                                                      "Medium",
                                                      "High")))
```

## Data Summaries

To begin, let's take a look at the data structure:

```{r data glance}
head(lydemap::lyde)
```

Each data point contains information on its source and specific dataset of origin ("survey"). The data is organized by year (specified as both calendar "year" and "bio_year", running from May 1st to April 30th), coordinates, and state. Additional columns define whether slf was found (even as an anecdotal individual record) at that location ("slf_present"), whether an established population was found there ("slf_established"), and what population density was ("slf_density"). For additional information on the variables included, please consult the help file associated with the data by typing `?lyde` in the RStudio console. A Metadata file can also be found in the compressed folder `lyde_data.zip` contained in `download_data/`.

The package function `data_summary` breaks the data down into a quick summary, with data organized by different axes. We can take a look at the data split across year and States. It's important to notice that the data is arranged yearly according to the **biological year** of SLF, and not calendar year. This allows for the appropriate inclusion of egg masses discovered during the winter months which were laid during the previous calendar year's summer/fall. 

```{r data by year and state, results = "asis", fig.cap = "Data by Biological year and State"}
# data by Year and State
knitr::kable(table(lyde$state, lyde$bio_year, useNA = "ifany"))
```


## Maps of the Spread of SLF

Two functions allow the user to plot the data: `map_spread()` and `map_yearly`.

### `map_yearly()`

The first function produces a snapshot of the SLF spread in the United States, with reference to the sampling effort associated with surveying the spread. Surveyed finding an established population are plotted on the map as filled circles, color coded by the year of first discovery. Surveys finding no established population are plotted as grey crosses.

```{r single map, fig.cap = "Map of SLF records", fig.height = 8, fig.width = 7.5}
# We can now proceed to plotting
map_spread()
```

The function can be customized to show the data gridded at higher spatial resolution (10k^2^), by setting the option `resolution` to "10k".

```{r map at 10k, fig.cap = "Map of SLF records at 10km resolution", fig.height = 8, fig.width = 7.5}
map_spread(resolution = "10k")
```

In addition, the user can specify which area of the United States should be mapped, by setting the `zoom` variable to "custom" and specifying the boundaries of the mapped area through `xlim_coord` (longitude) and `ylim_coord` (latitude).

```{r map at 10k, fig.cap = "Map of SLF records at 10km resolution", fig.height = 8, fig.width = 7.5}
map_spread(resolution = "1k",
           zoom = "custom",
           xlim_coord = c(-78, -74),
           ylim_coord = c(38, 42))
```

### `map_yearly()`

The second function allows the user to visualize the progression of SLF establishment, and the collection of surveys, over time. Note that the data here is not cumulative, meaning only data from a given year is shown in a panel of the figure.

```{r density map, fig.cap = "Map of population density by Year", fig.height = 7, fig.width=7}
# running year-specific map
# map_yearly()
```





